---
title: 'HW3 intermediate R'
output:
  pdf_document: default
  html_document:
    keep_md: yes
---
To hand in: Rmd and knitted pdf with the answers to the exercises.

Full score = 30 points + 6 overall points


# Introduction
HW3 continues with the data set from HW1, the functions for logit-transformations, and the `logit_min001`-transformed `Median` values. Those are included below. 

```{r setup, include = TRUE}
library(tidyverse)
library(readxl)
library(modelr)
library(broom)
library(gghighlight)
```

```{r, include = TRUE}
est0 <- read_excel("Table_Model-based_estimates_Countries_2019.xlsx", 
                   skip = 4, sheet = "FP Indicators")

logit <- function(x){
  log(x/(1-x))
}

logit_min001 <- function(x){
  ifelse(x < 0.01, logit(0.01), logit(x))
}

est <- 
  est0 %>%
  rename(iso = "ISO code", bounds = "Median estimate and uncertainty intervals",
         mar = "Marital status", country = "Country or area") %>%
  filter(Indicator == "CPmod",  mar == "MW") %>%
  pivot_wider(names_from = bounds, values_from = Percentage) %>%
  mutate(logit_cp = logit_min001(Median/100))

est
```

# Question 1 (10 pts)
For all countries, plot `logit_min001`-transformed `Median` (the `logit_cp` as above) against time. Facet wrap by the value of `Median` in the year 2030, with `high` defined as countries with prevalence greater than 50\%, `low` otherwise. The example plot is shown below.

Tip: check Biostats 597D intro2R if you need some inspiration on how to categorize the countries.

```{r, include = TRUE}
my_plot_est <- est %>% 
  group_by(country) %>% 
  mutate("type" = ifelse((Median[Year%in%2030] > 50),"high","low")) %>%
  ggplot(mapping = aes(x = Year, y = logit_cp, group = country)) + geom_line() + facet_wrap(~type)
my_plot_est
```

# Question 2 (10 pts)
For the United States of America, plot `logit_min001`-transformed `Median` (the `logit_cp` as above) over time and add a fitted regression line based on a linear regression model with the `logit_min001`-transformed `Median` as outcome variance, and predictor Year. Also plot the residuals of that model over time. Example plots are as follows:

```{r, include = TRUE}
my_plot <- est %>% 
  filter(country == "United States of America")

my_plot_2 <- my_plot %>% lm(logit_cp~Year,data = .) %>% augment()

ggplot(data = my_plot_2, mapping = aes(x = Year,y = logit_cp)) +      geom_point() + ggtitle("United State of America") + 
  geom_line(mapping = aes(x = Year, y = .fitted), color = "blue")
  
```


```{r, include = TRUE}
my_resid_plotest <- my_plot_2 
ggplot(data = my_resid_plotest, mapping = aes(x = Year, y = .resid)) + geom_point() + ggtitle("United State of America") + ylab("Residuals")
  
```


# Question 3(10 pts)
Fit a regression model to the data in each country and obtain the residuals. Then make the plot below that shows the residuals over time for each country in the data set. 

Tips: You can write your own for-loop if you want to practice more with the material you learned so far. Or you can consider using `group_by()` and `do()`, as explained in the first part of the optional Modeling lecture slides ("3 modeling2_optional.pdf") on Google Drive. 


```{r, include = TRUE}
my_plot_all <- est %>% group_by(country) %>% 
  do(lm(logit_cp~Year,data = .) %>% augment())
ggplot(data = my_plot_all)+
  geom_line(aes(Year,.resid,group = country),alpha = 0.2)
```


Extra credit (5 pts): highlight the USA in blue

```{r, include = TRUE}
my_plot_all_2 <- ggplot(data = my_plot_all)+
  geom_line(aes(Year,.resid,group = country),alpha = 0.2)+
  gghighlight(country == "United State of America")
my_plot_all_2
```





